\documentclass[a4paper,11pt]{article}

\usepackage[hidelinks]{hyperref}
\usepackage{geometry}
\geometry{margin=2.5cm}

\begin{document}

\begin{center}
{\Large \textbf{List of Publications}}
\end{center}

\vspace{1em}

\section*{Peer-Reviewed Journal Articles}

\begin{enumerate}
\item \textbf{Q. H. Nguyen}, O. Sadki, H. Rafaralahy, M. Haddad, A. Zemouche,  
\emph{Cyberattack Detection by Using a Discrete-Time Model-Based Unknown Input Observer},  
IEEE Control Systems Letters, vol. 8, pp. 856--861, 2024.
\end{enumerate}

\section*{International Conference Papers}

\begin{enumerate}
\item \textbf{Q. H. Nguyen}, S. Meng, M. Haddad, H. Rafaralahy, A. Zemouche,  
\emph{Neural Observer-Based Learning for Robust State Estimation},  
International Conference on Electrical and Computer Engineering (ICEECE), Madagascar, 2025.

\item S. Bougherara, \textbf{Q. H. Nguyen}, M. Haddad, H. Rafaralahy, A. Zemouche,  
\emph{Application of Discrete-Time Unknown Input Observers for Cyberattack Detection in Connected CACC Vehicles},  
ICEECE, Madagascar, 2025.

\item S. Meng, \textbf{Q. H. Nguyen}, A. Zemouche, F. Meng, F. Zhang,  
\emph{Distributed High-Gain Observer for Nonlinear Connected Autonomous Vehicle},  
American Control Conference (ACC), 2025.

\item O. Sadki, \textbf{Q. H. Nguyen}, A. Zemouche, H. Rafaralahy, M. Haddad,  
\emph{Cyberattack Estimation in Intelligent Transportation System by Using an Adaptive High Order Sliding Mode Differentiator},  
33rd Mediterranean Conference on Control and Automation (MED), 2025.

\item \textbf{Q. H. Nguyen}, O. Sadki, H. Rafaralahy, M. Haddad, A. Zemouche,  
\emph{Cyberattack Detection by Using a Discrete-Time Model-Based Unknown Input Observer},  
IEEE Conference on Decision and Control (CDC), 2024.
\end{enumerate}

\section*{Submitted / In Preparation}

\begin{enumerate}
\item \textbf{Q. H. Nguyen}, S. Meng, M. Haddad, H. Rafaralahy, A. Zemouche,  
\emph{Resilient Trust--Aware Distributed Observer Design for Connected Vehicle Platoons},  
23rd IFAC World Congress, 2026 (submitted).

\item S. Meng, \textbf{Q. H. Nguyen}, M. Alma, A. Zemouche,  M. Haddad, 
\emph{Distributed Unknown Input Observer for Vehicle Platoons Under Sensor Faults.} 
{23rd IFAC World Congress, 2026 (Submitted)}

\item \textbf{Q. H. Nguyen}, S. Meng, M. Haddad, H. Rafaralahy, A. Zemouche,  
\emph{Control-Oriented Trust-Aware Observer Design for Connected Vehicle Platoons},  
IEEE Transactions on Intelligent Transportation Systems (in preparation).
\end{enumerate}

\section*{Poster Presentation}

\begin{enumerate}
\item \textbf{Q. H. Nguyen}, O. Sadki, H. Rafaralahy, M. Haddad, A. Zemouche,  
\emph{Cyberattack Detection by Using a Discrete-Time Model-Based Unknown Input Observer},  
American Control Conference (ACC), 2024.
\end{enumerate}

\end{document}

% \Myprod{\textbf{Q. H. Nguyen}, S. Meng, M. Haddad, H. Rafaralahy, A. Zemouche}
% {Control-Oriented Trust-Aware Observer Design for Connected Vehicle Platoons.} 
% {IEEE Transactions on Intelligent Transportation Systems (Preparation)}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section*{Poster Presentations}

% \Myprod{\textbf{Q. H. Nguyen}, O. Sadki, H. Rafaralahy, M. Haddad, A. Zemouche}
% {Cyberattack Detection by Using a Discrete-Time Model-Based Unknown Input Observer}{(ACC), 2024}
% \end{document}