\chapter{Remerciements}

Many people have provided invaluable support throughout my doctoral journey, including my supervisors, colleagues, family, and my girlfriend.

First and foremost, I would like to express my sincere and deepest gratitude to my supervisors, Prof. Ali Zemouche, Prof. Rafaralahy, and Mr. Madjid Haddad, for giving me the opportunity to pursue my PhD research. I am profoundly grateful for their continuous guidance, encouragement, and trust. Their insightful advice, valuable discussions, and constant support—both scientific and personal—have been essential to the completion of this dissertation. Without their help, this work would not have reached its present form.

I am also extremely grateful to all my colleagues—Hasni, Shengya, Echrak, Hichem, Oussama, Sesabil, Tansuh, Viet, Melissa, and Li Qi—for their cooperation, constructive comments, and wise advice throughout my PhD studies. 
From my very first day in the laboratory, through my first scientific publication, they were always willing to offer help and support. 
When I led a team project for the first time, 
despite my limited experience in a leadership role, 
these colleagues I supervised during my internship demonstrated remarkable patience and understanding. 
Rather than blaming my mistakes, 
they supported me with tolerance and encouragement, 
allowing me to grow both professionally and personally.

Similarly, during my first teaching experiences, 
I faced significant challenges due to fear, shyness, and 
the difficulty of teaching in French, which is not my native language.
Their continuous support and encouragement played a crucial role in helping me overcome these obstacles and build confidence in myself.



Last but not least, I would like to express my deepest gratitude to my family—my uncle and aunt, my parents, and my girlfriend—for their unconditional love, patience, and support. I am especially thankful to my uncle and aunt, who have continuously supported me both financially and spiritually throughout my entire academic journey. Their encouragement and sacrifices have been a constant source of motivation and have greatly contributed to my pursuit of higher academic achievements.